\input opmac
\input diplomka_style
% \draft

\citaceAPA

\title={Algorithms for automated logistics}

% \margins/1 a4 (35,40,40,50)mm
\margins/1 a4 (40,25,25,25)mm

\emergencystretch=1em

\fixmnotes\left

% \def\pozn#1{}

\footline={}

\begingroup \fourteenrm
\centerline{\pozn{Gen.: \the\day. \the\month. \the\year}Charles University in Prague}
\medskip
\centerline{Faculty of Mathematics and Physics}
\vfill
\centerline{\sixteenbf MASTER THESIS}
\vfill
\centerline{\picw=64mm \inspic mfflogo.pdf }
\vfill
\vskip5mm
\centerline{\eighteenbf Bc. Michal Tuláček}
\vskip15mm
\centerline{\eighteenbf \the\title}
\vfill
\centerline{Department of Theoretical Computer Science and Mathematical Logic}
\vfill

\setbox0=\vbox{
\+Supervisor of the master thesis: &prof. RNDr. Roman Barták, Ph.D.\cr
\vskip5mm
\+\hfill Study programme: &Computer science\cr
\vskip5mm
\+\hfill Specialization: &Theoretical Computer Science\cr
}
\centerline{\box0}
\vfill
\centerline{Prague 2014}
\endgroup
\eject

\noindent I wish to thank my supervisor, prof. Roman Barták, for his invaluable advice and help. I am grateful for the experimental data provided by the Sunnysoft company, especially, my thanks go to Ing. David
Šilhan and Ing. Martin Strouhal for their support and possibility to test the system in a real environment. My greatest thank-you belongs to you, my sweet Martina. Without your help, support and understanding,
 I would have never finished this thesis.

\vfill

\eject

\null
\vfill

\noindent I declare that I carried out this master thesis independently, and only with the cited 
sources, literature and other professional sources.

\vskip5mm

\noindent I understand that my work relates to the rights and obligations under the Act No.
121\discretionary{/}{/}{/}2000~Sb., the Copyright Act, as amended, in particular the fact that the Charles
University in Prague has the right to conclude a license agreement on the use of this
work as a school work pursuant to subsection 60 (1) of the Copyright Act.

\vskip5mm

\noindent Prague, 31\st July 2014

\vskip15mm

\eject
\parskip=6pt
\bgroup\parindent=0pt
{\chyph
{\bf Název práce:} Algoritmy pro automatizovanou logistiku

{\bf Autor:} Michal Tuláček

{\bf Katedra:} Katedra teoretické informatiky a matematické logiky

{\bf Vedoucí diplomové práce:} prof. RNDr. Roman Barták, Ph.D.

{\bf Abstrakt:} Tématem práce je automatická logistika maloobchodního řetězce. Získané poznatky jsou aplikovány na reálný problém maloobchodní společnosti provozující elektronický obchod a několik poboček 
v různých městech. V úvodní části je nejprve důkladně prostudován reálný problém řešený touto společností. Následně jsou prezentované současné trendy v oblasti řešení tohoto problému, zejména s 
přihlédnutím k řešením vycházejících z problému směrování vozidel a z problémů s omezujícími podmínkami. Navržené řešení je podrobeno testům vycházejícím z dat ze skutečného provozu.

{\bf Klíčová slova:} automatická logistika, plánování, rozvrhování
}

\vfil

{\bf Title:} \the\title

{\bf Author:} Michal Tuláček

{\bf Department:} Department of Theoretical Computer Science and Mathematical Logic

{\bf Supervisor:} prof. RNDr. Roman Barták, Ph.D.

{\bf Abstract:} The topic of this thesis is called algorithms for automated logistics. The gathered findings are applied on a real issue of a retail company running an electronic shop and a few branches 
in several cities. In the introduction part, there is a thorough description of an actual problem that is being dealt with by this company. Consequently, contemporary trends in the area of 
problem-solving of this issue is presented, especially regarding the answers to vehicle routings and issues with limiting conditions. The offered solution is subjected to tests arising from 
the data collected from the real-time service.

{\bf Keywords:} automated logistics, planning, scheduling
    
\vfil\egroup
\eject

{\chapfont\noindent Contents}
\vskip5mm
\maketoc

\vfill

\eject

\footline={\hss\folio\hss}
\def\zahlavi{\noindent{\it \kern3pt\headrule\kern3pt \the\title}}
\input chaps/introduction-new
\input chaps/problem-analysis
\input chaps/models
% \input chaps/strategies
\input chaps/problem-model
% \input chaps/basic-algorithm
% \input chaps/problem-solution
\input chaps/benchmarking
%\input chaps/discussion
\input chaps/conclusion

\vfill\eject
{\sdef{mt:chap:en}{}\nonum\chap Bibliography\par}
\makebib

\global\chapnum=0
\input appendices/abbrev
\input appendices/content-cd
\input appendices/implementation
\input appendices/simulator
\input appendices/data-files

\bye
