\newpage

\label[chap:benchmarking]
\chap Benchmarking

\pozn{Zkontrolovano 15.7.}To evaluate the model proposed in the previous chapter, we compare it with the algorithm described in chapter \ref[chap:analysis]. We perform tests on real data collected during fiscal year 07/2013---06/2014.

\sec A simulator for automated logistics

As we described in chapter \ref[chap:analysis], the whole process is controlled by documents. Therefore, based on the data in ERP, we are able to establish the $InStock$ values for each store in any
given moment. Moreover, we have the demands represented by shipment orders and also warehouse receipts which indicate if and when there was an increase in the amount of goods in a given store.

In order to test our algorithm, we created a simulator which generates demands for the planner and applies the generated plan.

\begitems \style N
* {\em Intialization}. For a given date $d$, the simulator sets $Availability_{g,s} = InStock_{g,s}$ based on the history of stock supplies. Then it prepares a sequence of demands ${\bf D}$ and 
a sequence of store replenishments ${\bf R}$, both ordered by date, where the date of the first element of both sequences is after $d$. We introduce a sequence of shedules ${\bf Sch} = \emptyset$.
* {\em Pre-planning phase}. In this phase, the simulator sets $d'=d$ and increases $d$ by a predefined step. Then it constructs a sequence ${\bf D'} \subset {\bf D}$ of demands, which occurred since $d'$.
Next, it updates an $Availability_{g,s}$ with replenishments from ${\bf R}$ which occurred since $d'$ and removes them from ${\bf R}$. Similarly, it processes schedules in ${\bf Sch}$. For each schedule
$sch \in {\bf Sch}$ which is planned between $d'$ and $d$, the simulator lowers the corresponding $OnTheWay_{g,s}$ and increases $Availability_{g,s}$. Then it computes a set of shortest paths ${\bf P}$ 
between the stores, with shipment time $d$.
* {\em Planning}. The simulator runs the selected planner algorithm with ${\bf D'}$ and ${\bf P}$ as an input, producing a set of plans ${\bf Pl}$.
* {\em Update phase}. Each plan $pl \in {\bf Pl}$ contains an information about the origin store $s_o$, the destination store $s_d$, goods to be transferred $s_o$, its $amount$ and a reference to 
a demand $d$, which is resolved by the plan. According to the plan, the simulator decreases the requested amount of goods $g$ in $d$. If the demand $d$ is completely resolved, i.e. it does not demand
any goods, it is removed from ${\bf D}$. Then it decreases $Availability_{g,s_o}$ and increases $OnTheWay_{g,s_d}$ by $amount$. Finally, it adds a new schedule to ${\bf Sch}$ for transfer of $amount$ 
goods $g$ to store $s_d$ with a scheduled time based on the shortest path in ${\bf T}$ between $s_o$ and $s_d$. Afterwards, it continues with phase 2).
\enditems

\sec Description of the benchmarks

We test the model on data from 1\st November to 24\th December 2013. Due to the increased pre-Christmas demand rate, this period is the most challenging part of the year. We run the 
simulations with the CSP model, described in Chapter \ref[chap:model], Na\"{\i}ve algorithm described in section \ref[secc:analysis:naive] and Ford-Fulkerson based algorithm in section
\ref[secc:analysis:ford-fulkerson].

For each planning algorithm, we run a simulation with a 1-minute, 5-minute and 30-minute-long time step. Each of the configurations is run either with or without the low-priority demands.

Since the CSP model has two parameters, $\alpha$ and $\beta$, we will run simulations with low-priority demands with various values of the parameters.

We measure the following quantities:

\begitems
* Number of resolved demands.
* Total time needed to completely resolve the demands.
* Number of unresolved demands.
* For each store a number of days when the store was overloaded due to exceeding the capacity.
\enditems

We do not compare the algorithms according to the running time, since it is not a critical issue for the problem. However, the CSP solver has a limit in finding the optimal solution. If it does not
find a solution within one minute of the running time, the problem is considered as not being feasible.

\secc Configuration of stores and deliveries

The configuration of stores and delivery routes used in the benchmarks is presented on Figure \ref[fig:benchmarking:deliveries]. There are six stores -- central store (C), e-shop (E), Kovanecká 
street (K), passage Grossmann (G), Brno (B) and 
Plzeň (P). Three of them (central store, e-shop and Kovanecká street) are in one building. Therefore, the delivery is instant and can occur during the whole business day. The Grossmann store is 
located in Prague and its delivery is maintained by a courier company. Therefore, it is not instant but the goods is still delivered in the same day as it is shipped. It is shipped and delivered at a regular
fixed time. Finally, other routes are inter-city routes which occur only once a day at a fixed time. The shipment and delivery times are specific for a particular store.

\begfigure
\centerline{\inspic figs/benchmarking-deliveries-1.pdf }\nobreak\medskip
\label[fig:benchmarking:deliveries]\caption/f The stores and delivery routes used in benchmarking. The row {\em s} represents the conditions for a shipment and the row {\em d} contains the duration
of a transfer along the path and a corresponding delivery time.
\endfigure

\secc Low-priority demands

The low priority demands are generated once a day between 6:00 and 6:59. For each goods $g$ the simulator computes the expected available amount of goods after resolving all of the demands. It uses
the following formula:

$$ExpectedTotalAvailability_g = \sum_{s \in {\bf S}}{InStock_{g,s}} - \sum_{d \in {\bf D}}{Demanded_{g,d}}, \eqmark$$

\noindent where $Demanded_{g,s}$ means the amount of goods $g$ demanded by the demand~$d$. Afterwards, we compute desired amount on the stores. Since we want to balance amounts of goods equally, we use the 
following formula

$$DesiredAmount_g = \left\lceil ExpectedTotalAvailability_g \over |{\bf S}| \right\rceil, \eqmark$$

\noindent and generate the low-priority demand $d$ for each store with 

$$Demanded_{g,d} = DesiredAmount_g.$$

\secc Shortest paths

To compute a set of paths ${\bf P}$ between the stores, we use a Dijkstra's algorithm \cite[Dijkstra1959] on a directed graph displayed on Figure \ref[fig:benchmarking:deliveries]. The paths
are computed for each run once again, since the shortest path depends on a current simulation's date and time. The lengths of paths are computed using the Algorithm \ref[alg:algorithm:delivery-time].

\sec Experimental results


\sec Discussion

